---
title: "&nbsp;v2.0"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    keep_md: no
    includes:
        before_body: histomx_header.html
        after_body: histomx_footer.html
params:
  rcc_file: 'missing-filename'
  rna_file: ''
  patient_file: ''
---

<style type="text/css">
  body{font-size: 12pt;}
</style>

```{r, include=FALSE, eval=FALSE}
## to run manually
newRCC='../test_files/test.RCC'
newPatient='../test_files/patient-test.json'
rmarkdown::render('../scripts/histomx_kidney.Rmd')
## convert output html to paged PDF 
system('wkhtmltopdf --javascript-delay 1 ~/Desktop/histomx_report.html ~/Desktop/histomx_report.pdf')
```

```{r setup, include=FALSE, echo=FALSE}
options(rgl.useNULL=TRUE)
## set paths in all code chunks are relative to this root.dir
rdir=getwd()
knitr::opts_knit$set(rdir)

library(htmltools, quietly=TRUE)
library(ggplot2, quietly=TRUE)
theme_set(theme_minimal())
library(gridExtra, quietly=TRUE)
library(knitr, quietly=TRUE)
opts_knit$set(eval.after = "fig.cap")
library(kableExtra, quietly=TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

***
## Molecular Pathology Platform
> Inserm U970, 56 rue Leblanc 75015 Paris, France  
> **Team leader:** Pr Alexandre Loupy MD/PhD  
**Project managers:** Fariza Mezine & Dina Zielinski PhD  
**Contact:** jessy.dagobert@inserm.fr / +33-1-53-98-80-85

```{r, echo=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
source('../scripts/BHOTpred.R')
```

```{r, echo=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
## import parameters (list of input files)
## see render-rmarkdown.sh to run markdown with new RCC as input
newPatient <- params$patient_file
newRNA <- params$rna_file
newRCC <- params$rcc_file

newPath <- normalizePath(dirname(newRCC))
dir.create(paste0(newPath, '/histomx_results/'), recursive=TRUE, showWarnings=FALSE)
outPath=paste0(newPath, '/histomx_results/')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## the following functions output all files to a directory called 'histomx_results' in the input file path
newPreds <- BHOTpred(newRCC, out_path=outPath, save_files=TRUE, norm_method="separate", output_id="sample_id")
newID <- newPreds$new_scores$ID
```

***
## Sample info:
<div style="margin-bottom:40px;">
```{r, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE}

## check if sample info exists; check file type (json or txt) and import
if(!is.na(file.info(newPatient)$size)) {
  
	if (tools::file_ext(newPatient) == "json") {
		jfile <- fromJSON(file=newPatient)
		jfile <- jsonlite:::null_to_na(jfile) # convert NULL to NA
		pt_df <- as.data.frame(jfile, check.names=F, fix.empty.names=F)
		pt_df <- data.frame(V1=rownames(t(pt_df)), V2=t(pt_df), check.names=F)
	} else if (tools::file_ext(newPatient) == "txt") {
		pt_df <- read.table(newPatient, sep='\t', na.strings=c("", "NA"), encoding="UTF-8")
	} else {
		stop("Sample info file must be a .json or tab delimited .txt file")
	}	

	## if all vars in imported file are null, display empty data frame
	if (nrow(pt_df) == 0) {
		pt_df <- data.frame('Platform ID'=newPreds$new_scores$ID,
                        'Report date'=Sys.Date(), check.names=FALSE)
		pt_df <- data.frame(V1=rownames(t(pt_df)), V2=t(pt_df)[,1])
	}
	## exclude empty cells
	pt_df <- pt_df %>% dplyr::filter_all(all_vars(!is.na(.)))
      
} else if (exists('pt_df')==FALSE) {
	## if no file, display empty data frame
	pt_df <- data.frame('Platform ID'=newPreds$new_scores$ID,
                        'Report date'=Sys.Date(), check.names=FALSE)
	pt_df <- data.frame(V1=rownames(t(pt_df)), V2=t(pt_df)[,1])
}

## format and output patient info table
if(nrow(pt_df) %% 2 == 0) {
    pt_df <- cbind(pt_df[1:(nrow(pt_df)/2),], pt_df[(nrow(pt_df)/2+1):nrow(pt_df),])
    kable(pt_df, row.names=FALSE, col.names=rep("", ncol(pt_df)), align='c', format="html") %>%
          row_spec(1:nrow(pt_df), background="white") %>%
          column_spec(c(1,3), bold=TRUE, background="whitesmoke") %>% kable_styling()
} else {
    ## todo: set 1st row as single row, remaining as double
    kable(pt_df, row.names=FALSE, col.names=rep("", ncol(pt_df)), align='c', format="html") %>%
    row_spec(1:nrow(pt_df), background="white") %>%
    column_spec(1, bold=TRUE, background="whitesmoke") %>% kable_styling()
}

```
</div>

\newpage

***
## Gene expression based probabilities:
```{r, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE}

## molecular score table
ms_new <- data.frame(t(newPreds$new_scores[,c("amr", "tcmr", "normal")]), check.names=FALSE)
ms_new$Diagnosis <- rownames(ms_new)

## confidence intervals
ci_tab <- data.frame(newPreds$new_scores[,grep("_ci", names(newPreds$new_scores))])
ci_new <- data.frame(amr=paste(ci_tab$amr_lwr_ci, "-", ci_tab$amr_upr_ci),
		     tcmr=paste(ci_tab$tcmr_lwr_ci, "-", ci_tab$tcmr_upr_ci),
		     normal=paste(ci_tab$normal_lwr_ci, "-", ci_tab$normal_upr_ci))
ci_new <- data.frame(CI=t(ci_new))
ci_new$Diagnosis <- rownames(ci_new)

## reference set quantiles
ms_ref <- newPreds$ref_scores[,c("Dx", "amr", "tcmr", "normal", "Q")]
 
## q1 and q3 (can also add IQR, min, max)
tab <- data.frame(matrix(nrow=dim(ms_new)[1], ncol=2), check.names=FALSE)
colnames(tab) <- c("No rejection Q1", "No rejection Q3")
rownames(tab) <- rownames(ms_new)

## rows=molecular scores; columns=histo diagnosis

## reference biopsy AMR molecular score ranges
tab[rownames(tab)=="amr","No rejection Q1"] <- ms_ref[ms_ref$Dx=="NONREJECTION" & ms_ref$Q=="25%","amr"]
tab[rownames(tab)=="amr","No rejection Q3"] <- ms_ref[ms_ref$Dx=="NONREJECTION" & ms_ref$Q=="75%","amr"]

## reference biopsy TCMR molecular score ranges
tab[rownames(tab)=="tcmr","No rejection Q1"] <- ms_ref[ms_ref$Dx=="NONREJECTION" & ms_ref$Q=="25%","tcmr"]
tab[rownames(tab)=="tcmr","No rejection Q3"] <- ms_ref[ms_ref$Dx=="NONREJECTION" & ms_ref$Q=="75%","tcmr"]

## reference biopsy non-rejection molecular score ranges
tab[rownames(tab)=="normal","No rejection Q1"] <- ms_ref[ms_ref$Dx=="NONREJECTION" & ms_ref$Q=="25%","tcmr"]
tab[rownames(tab)=="normal","No rejection Q3"] <- ms_ref[ms_ref$Dx=="NONREJECTION" & ms_ref$Q=="75%","tcmr"]

tab$'Normal range' <- paste(tab$`No rejection Q1`, "-", tab$`No rejection Q3`)
tab$'Diagnosis' <- rownames(tab)
tab <- merge(tab, ci_new, by="Diagnosis")

dx_tab <- merge(ms_new, tab[,c("Diagnosis", "CI", "Normal range")], by="Diagnosis")
colnames(dx_tab)[colnames(dx_tab)=="CI"]<-"Sample range"

dx_order <- c("amr", "tcmr", "normal")
dx_tab <- dx_tab %>% dplyr::slice(match(dx_order, Diagnosis))
dx_tab$Diagnosis[dx_tab$Diagnosis=="amr"]<-"AMR"
dx_tab$Diagnosis[dx_tab$Diagnosis=="tcmr"]<-"TCMR"
dx_tab <- dx_tab[!dx_tab$Diagnosis %in% "normal",] #exclude NONREJECTION score
#dx_tab <- dx_tab %>% relocate('Diagnosis', .before = colnames(dx_tab)[1])

#TODO: scale probabilities relative to normal biopsies (eg. +/- SD)
dx_tab$Interpretation <- ifelse(dx_tab[,newID]<=10,"unlikely",
                                       ifelse(dx_tab[,newID]>10 & dx_tab[,newID]<=20,"low probability",
                                              ifelse(dx_tab[,newID]>20 & dx_tab[,newID]<=40,"low/moderate probability",
                                                     ifelse(dx_tab[,newID]>40 & dx_tab[,newID]<=60,"moderate probability",
                                                           ifelse(dx_tab[,newID]>60 & dx_tab[,newID]<=80,"high probability",
                                                                  ifelse(dx_tab[,newID]>80,"very high probability","NA"))))))

kable(dx_tab, row.names=FALSE, align='c', caption="", format="html") %>%
    column_spec(1, bold=TRUE, color="black") %>%
    column_spec(c(2,3), bold=TRUE, color="dodgerblue") %>%
    row_spec(c(1,2), background="white")

```
Diagnosis based scores represent the probability (0-100%) that a biopsy has an expression profile similar to reference samples with a given histology based rejection diagnosis (AMR or TCMR).  
'Sample range' is the confidence interval for the predicted probabilities of the tested sample.  
'Normal range' is the middle 50% of scores in biopsies with no rejection or injury diagnosis.

***
## Principal Components Analysis (PCA) of molecular scores:
<div style="margin-bottom:50px;">
```{r, echo=FALSE, warning=FALSE, comment=FALSE, results='asis', fig.align='center', fig.cap=''}

#gridExtra::grid.arrange(newPreds$pca_1_2, newPreds$pca_2_3, ncol=2)
newPreds$pca_1_2
#newPreds$pca_2_3

```
Samples are labeled by histologic diagnosis (NRKI=non-rejection kidney injury). The tested sample is the yellow diamond.
</div>

\newpage

***
### Histologic Diagnosis of 25 Nearest Neighbors
```{r, echo=FALSE, warning=FALSE, results='asis', eval=TRUE}
newPreds$knn_dx$Dx <- rownames(newPreds$knn_dx)
newPreds$knn_dx <- newPreds$knn_dx[,c("Dx", "Total", "Percent")]
colnames(newPreds$knn_dx) <- c("Diagnosis", "Total", "Percent")

kable(newPreds$knn_dx, row.names=FALSE, align='c', caption="", format="html") %>%
    column_spec(1, bold=TRUE, background="white", color="black") %>%
    column_spec(c(2,3), bold=TRUE, color="dodgerblue", background="white")
```

***
### Clinician signout
```{r, echo=FALSE, warning=FALSE, include=TRUE}
sign_df <- data.frame(Name=">", Signature="")

kable(sign_df, align='c', format="html") %>% kableExtra::kable_styling(position="center") %>%
  row_spec(1, background="white", align='l')
```
> **_Comments:_**  
>  &nbsp;    
>  &nbsp;  
>  &nbsp;  
>  &nbsp;    
>  &nbsp;  
>  &nbsp;    
>  &nbsp;  

\newpage

***
## Technical details
***

### RNA quality
```{r, echo=FALSE, warning=FALSE, include=TRUE}

if(!is.na(file.info(newRNA)$size)) {
  
  	if (tools::file_ext(newRNA) == "json") {
		jfile <- fromJSON(file=newRNA)
		jfile <- jsonlite:::null_to_na(jfile) # convert NULL to NA
		rna_df <- as.data.frame(jfile, check.names=F, fix.empty.names=F)
		rna_df <- data.frame(V1=colnames(rna_df), V2=t(rna_df))
	} else if (tools::file_ext(newRNA) == "txt") {
		rna_df <- read.table(newRNA, check.names=F, sep='\t')
	} else {
		stop("Sample info file must be a .json or tab delimited .txt file")
	}	

	## create data frame if imported variables all null
	if (nrow(rna_df) == 0) {
		rna_df <- data.frame(V1=c("platform id", "center_id", "260/280 ratio", "260/230 ratio", "concentration", "units"),
				     V2=c(newPreds$new_scores$ID, NA, NA, NA, NA, "ng/ul"))
	}
	
	## auto-fill units if missing
	if(!c("units") %in% rna_df[,1]) {
		cat("Adding a column for units (ng/ul)")
		rna_df <- rbind(rna_df, data.frame(V1="units", V2="ng/ul"))
	}
			
	## transpose and use first column as header
	if(nrow(rna_df) > 2) {
		rna_df <- setNames(data.frame(t(rna_df[,-1])), rna_df[,1])
	}

	## if units (ng/ul) included in RNA concentration strip; else convert to numeric
	#if(length(strsplit(rna_df$concentration, " ")[[1]]) > 1) {
	#	rna_df$concentration <- as.numeric(strsplit(rna_df$concentration, " ")[[1]][1])
	#} else {
	#	rna_df$concentration <- as.numeric(rna_df$concentration)
	#}

	## move concentration and units to end
	rna_df <- rna_df %>% relocate(concentration, .after=last_col())
	rna_df <- rna_df %>% relocate(units, .after=last_col())

	## apply threshold based text color and add back units
	if (!is.na(rna_df$concentration)) {
		
		rna_df$concentration <- as.numeric(rna_df$concentration)
		
		if(rna_df$concentration >= 20) {
			rna_df$status <- "Passed"
			status_color="green"
		} else if(rna_df$concentration>=10 & rna_df$concentration<20) {
			rna_df$status <- "Caution"
			status_color="goldenrod"
		} else if(rna_df$concentration < 10) {
			rna_df$status <- "Warning"
			status_color="firebrick"
		}
		
		kable(rna_df, row.names=FALSE, align='c', caption="", format="html") %>%
		column_spec(1, bold=TRUE, color="dodgerblue") %>%
		column_spec(ncol(rna_df), bold=TRUE, color=status_color) %>%
		row_spec(1, background="white")
		
	} else {
		kable(rna_df, row.names=FALSE, align='c', caption="", format="html") %>%
		column_spec(1, bold=TRUE, color="dodgerblue") %>%
		column_spec(ncol(rna_df), bold=TRUE) %>%
		row_spec(1, background="white")
	}
	

} else {
	## display empty data frame
	rna_df <- data.frame(ID=newPreds$new_scores$ID, '260/280 ratio'="NA", '260/230 ratio'="NA",
                       concentration="NA", status="NA", check.names=FALSE)
  
	kable(rna_df, row.names=FALSE, align='c', caption="", format="html") %>%
	column_spec(1, bold=TRUE, color="dodgerblue") %>%
	column_spec(ncol(rna_df), bold=TRUE) %>%
	row_spec(1, background="white")
}

```
"Caution" status indicates a concentration within the minimum recommended 10-20ng/ul range  
"Warning" status indicates a concentration lower than the recommended minimum 10ng/ul

### NanoString run information
```{r, echo=FALSE, warning=FALSE, include=TRUE}

qcTab <- newPreds$qc_table
qcTab <- qcTab[!qcTab$variable %in% c("FovCount", "FovCounted", "ncgMean", "ncgSD", "FileVersion", "SoftwareVersion"),]
qcTab <- qcTab %>% dplyr::select(variable, everything())

qcTab <- cbind(qcTab[1:(nrow(qcTab)/2),], qcTab[(nrow(qcTab)/2+1):nrow(qcTab),])

kable(qcTab, row.names=FALSE, col.names=rep("", ncol(qcTab)), align='l', caption="", format="html") %>%
  column_spec(c(1,3), bold=TRUE, background="whitesmoke") %>%
  column_spec(c(2,4), background="white") %>% kable_styling()

#The lower the LOD the lower the expected level of background noise in the data
#https://www.nanostring.com/wp-content/uploads/2020/12/Gene_Expression_Data_Analysis_Guidelines.pdf
```
__FoV__=fields of view | __geoMean__=geometric mean | __PCL__=positive control linearity |
__LoD__=limit of detection | __BD__=binding density | __SNratio__=geoMean of HK genes:LoD

__Recommended QC thresholds:__    
Registered FoVs>75% | PCL>0.95 | BD 0.05-2.25 | 50% ENDO genes>LoD | POS_E counts>LoD | SNratio>1 | HK genes below LoD=0  
*Samples failing multiple QC flags and/or with >0 HK genes below LoD should be rerun.* 

### For research use only:
Models have been not been validated in clinical settings or trials. This report does not replace a standard diagnosis. Due to known technical variation, RNA later-fixed samples should be interpreted with caution.

```{r, echo=FALSE, include=TRUE}
logoIMG='../static/ptg-logo.png'

htmltools::img(src = knitr::image_uri(logoIMG), 
               alt = 'logo', 
               style="display: block; margin-left: auto; margin-right: auto;")
```

